Investigating Cross-Sectional Size Distribution of Randomly Distributed 3D Spheres

### 2020-05-24 Paper/main.tex
Mid way through results section

\begin{abstract}
  Text
\end{abstract}

@@

\section{Method}
Consider many spheres of different radii distributed randomly in space. For the
purposes of the numerical simulation, we will consider a cube of length 20. The
spheres were not allowed to overlap, which we ensured by placing the spheres such
that if one sphere has vector position $\overrightarrow{\textbf{a}}$ and radius
$r_{1}$ and the other has vector position $\overrightarrow{\textbf{b}}$ and
radius $r_{2}$, the spheres must have positions such that:
\begin{equation*}
|\overrightarrow{\textbf{a}} - \overrightarrow{\textbf{b}}| > (r_{1} + r_{2})
\end{equation*}
Random coordinates for a sphere are drawn from a uniform distribution and
checked against every existing sphere until a set of coordinates is drawn that
ensures no overlap, at which point the sphere is placed in space and the process
is repeated until all the necessary spheres have been placed.

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/box3d.pdf}
\caption{750 spheres of radius 0.7 distributed randomly in a cube of length 20,
  such that no sphere is touching or overlapping another sphere. A randomly
  oriented plane passes through the box and ``cuts'' through the
  spheres.}\label{fig:3dplot_plane}
\end{center}
\end{figure}

We then placed a plane into the cube, with a random orientation. The plane will
``cut'' through many of the spheres in the cube. This is shown in
Figure~\ref{fig:3dplot_plane}. Here we see 750 spheres, each with radius 0.7
distributed randomly in a cube of length 20. The plane is shown in blue, and can
be seen intersecting many of the spheres in the cube.

@@

\section{Results}
Text

% \begin{figure}[H]%[!ht]
% \begin{center}
% \includegraphics[width=0.45\textwidth]{./../Figures/box3d_noplane.pdf}
% \caption{Spheres of radius 0.7 distributed randomly in a cube of length 20, such
%   that no sphere is touching or overlapping another sphere.}
% \label{fig:3dplot}
% \end{center}
% \end{figure}

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_01.pdf}
\caption{Distribution of circle radii on plane for 750 spheres of radius $0.1$
  each. There are relatively few points since the spheres are small, so there is
  relatively small chance that they will be cut by the plane. So only a few
  circles are visible on the plane. The theoretical distribution is shown, with
  the experimental data matching this distribution, showing a slow increase in
  number or circles as radius increases.}\label{fig:size1}
\end{center}
\end{figure}

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_03.pdf}
\caption{Distribution of circle radii on plane for 750 spheres of radius $0.3$
  each. Compared to Figure~\ref{fig:size1}, there are many more experimental data
  points, since it is more likely that the plane will intersect these larger
  spheres. The experimental data more clearly resembles the theoretical
  distribution, showing a strong increase in number of circles as radius
  increases towards the maximum of 0.3.}\label{fig:size3}
\end{center}
\end{figure}

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_07.pdf}
\caption{Distribution of circle radii on plane for 750 spheres of radius $0.7$
  each. We do not observe much difference in how well the experimental data
  resembles the theoretical distribution between this figure and
  Figure~\ref{fig:size3}. With larger sized spheres, the simulation becomes harder
  to run as it is more difficult to randomly place spheres in a finite space
  without overlap.}\label{fig:size7}
\end{center}
\end{figure}

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_07_random.pdf}
\caption{Distribution of circle radii on plane for 750 spheres of radius $R$
  where $0.65 < R < 0.7$. The theoretical distribution is now different,
  although the experimental data still resembles this new distribution, showing
  a maximum around $R = 0.7$ and a sharp decrease after this. This is likely
  because there are more spheres with radii such that an overlap between $0$ and
  $0.7$ is possible, and relatively few spheres with radii such that a larger
  overlap is possible.}
\label{fig:random}
\end{center}
\end{figure}

@@

\begin{figure*}[ht!]
  \centering
  \subfloat[Noise 2 orders of magnitude smaller than radius]
  {\includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_100.pdf}
  \label{fig:2noise}}
  \centering
  \subfloat[Noise 1 order of magnitude smaller than radius]
  {\includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_10.pdf}
  \label{fig:1noise}}
  \caption{Distribution of circle radii on plane for 750 balls of radius $0.7$
    each. Each recorded radius has a random noise added to it in order to
    determine how sensitive the data is to experimental
    error. Figure~\ref{fig:2noise} shows the distribution for an added noise 2
    orders of magnitude smaller than the recorded radius, while
    Figure~\ref{fig:1noise} is the same for noise 1 order of magnitude smaller. As
    we can see, the experimental data very resembles the theoretical
    distribution when noise of 2 orders of magnitude smaller are added. But this
    is not true when noise of 1 order of magnitude smaller is added, implying
    the data is very sensitive to experimental data.}
  \label{fig:noiseplots}
\end{figure*}

\section{Discussion and Summary}
Text

@@

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%
\appendices{}
\section{Appendix Name}
Text

>>>>

\begin{abstract}
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
\end{abstract}

@@

\section{Method}

Consider many spheres of different radii distributed randomly in space. For the
purposes of the numerical simulation, we will consider a cube of length 20. The
spheres were not allowed to overlap, which we ensured by placing the spheres such
that if one sphere has vector position $\overrightarrow{\textbf{a}}$ and radius
$r_{1}$ and the other has vector position $\overrightarrow{\textbf{b}}$ and
radius $r_{2}$, the spheres must have positions such that:
\begin{equation*}
|\overrightarrow{\textbf{a}} - \overrightarrow{\textbf{b}}| > (r_{1} + r_{2})
\end{equation*}
Random coordinates for a sphere are drawn from a uniform distribution and
checked against every existing sphere until a set of coordinates is drawn that
ensures no overlap, at which point the sphere is placed in space and the process
is repeated until all the necessary spheres have been placed.

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/box3d.pdf}
\caption{$750$ spheres of radius 0.7 distributed randomly in a cube of length 20,
  such that no sphere is touching or overlapping another sphere. A randomly
  oriented plane passes through the box and ``cuts'' through the
  spheres.}\label{fig:3dplot_plane}
\end{center}
\end{figure}

We then placed a plane into the cube, with a random orientation. The plane will
``cut'' through many of the spheres in the cube. This is shown in
Figure~\ref{fig:3dplot_plane}. Here we see $750$ spheres, each with radius 0.7
distributed randomly in a cube of length 20. The plane is shown in blue, and can
be seen intersecting many of the spheres in the cube.

@@

\section{Results and Discussion}

We found that for both the case of the spheres of constant radius, and the case
of the spheres of varying radius, the experimental data reproduced a
distribution in the same shape as the theoretical distribution.

We first investigated the case where we placed $750$ spheres of radius $0.1$
each in a cube of length $20$. In this case we found that a plane in a random
orientation intersects very few spheres, since the spheres are very small. The
planar distribution is presented in Figure~\ref{fig:size1}.

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_01.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $0.1$
  each. There are relatively few points since the spheres are small, so there is
  relatively small chance that they will be cut by the plane. So only a few
  circles are visible on the plane. The theoretical distribution is shown, with
  the experimental data matching this distribution, showing a slow increase in
  number or circles as radius increases.}\label{fig:size1}
\end{center}
\end{figure}

We observe relatively few experimental data points, as there were relatively few
intersections with the plane. This was likely because the spheres' radii were
very small compared to box length and there were relatively few spheres. While
the theoretical distribution shows a strong increase near $0.1$, this is not
evident from the experimental data. We also see that there is a lot of variance
between the points so the slow rise evident in the theoretical distribution is
not clear in the experimental data. It is thus necessary that we have more
spheres that intersect with the plane, in order to obtain a larger data-set.

We found that increasing the spheres' size or placing more spheres in the box,
or making the box length smaller, are all good methods of increasing the number
of intersections we observe, as they all increase the likelihood of intersection
by increasing the ratio of sphere occupied space to total box space. We thus
next used $750$ spheres of radius $0.3$ each. The planar cross-sectional
distribution found is presented in Figure~\ref{fig:size3}.

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_03.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $0.3$
  each. Compared to Figure~\ref{fig:size1}, there are many more experimental data
  points, since it is more likely that the plane will intersect these larger
  spheres. The experimental data more clearly resembles the theoretical
  distribution, showing a strong increase in number of circles as radius
  increases towards the maximum of 0.3.}\label{fig:size3}
\end{center}
\end{figure}

In Figure~\ref{fig:size3} we see that there are clearly more experimental data
points. This is because more spheres had intersections with the plane. Compared
to Figure~\ref{fig:1noise}, the slow rise in the theoretical distribution is
more closely followed by the experimental data. However, we still encounter
significant variance and fluctuation in lower radius bins, even though the
theoretical data is a smooth gradual increase. This is likely due to the
relatively few number of circles populating these bins, causing greater relative
difference between neighbouring bins. One method to solve this would be to use
fewer bins, which would have the effect of ``summing and averaging''
neighbouring bins so the experimental data has a smoother trend. However, this
would also reduce the number of data points so it was not implemented in our
investigation.

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_07.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $0.7$
  each. We do not observe much difference in how well the experimental data
  resembles the theoretical distribution between this figure and
  Figure~\ref{fig:size3}. With larger sized spheres, the simulation becomes harder
  to run as it is more difficult to randomly place spheres in a finite space
  without overlap.}\label{fig:size7}
\end{center}
\end{figure}

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_07_random.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $R$
  where $0.65 < R < 0.75$. The theoretical distribution is now different,
  although the experimental data still resembles this new distribution, showing
  a maximum around $R = 0.7$ and a sharp decrease after this. This is likely
  because there are more spheres with radii such that an overlap between $0$ and
  $0.7$ is possible, and relatively few spheres with radii such that a larger
  overlap is possible.}
\label{fig:random}
\end{center}
\end{figure}

@@

\begin{figure*}[ht!]
  \centering
  \subfloat[Noise 2 orders of magnitude smaller than radius]
  {\includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_100.pdf}
  \label{fig:2noise}}
  \centering
  \subfloat[Noise 1 order of magnitude smaller than radius]
  {\includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_10.pdf}
  \label{fig:1noise}}
  \caption{Distribution of circle radii on plane for $750$ balls of radius $0.7$
    each. Each recorded radius has a random noise added to it in order to
    determine how sensitive the data is to experimental
    error. Figure~\ref{fig:2noise} shows the distribution for an added noise 2
    orders of magnitude smaller than the recorded radius, while
    Figure~\ref{fig:1noise} is the same for noise 1 order of magnitude smaller. As
    we can see, the experimental data very resembles the theoretical
    distribution when noise of 2 orders of magnitude smaller are added. But this
    is not true when noise of 1 order of magnitude smaller is added, implying
    the data is very sensitive to experimental data.}
  \label{fig:noiseplots}
\end{figure*}

\section{Conclusions}
Text

@@

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%
% \appendices{}
% \section{Appendix Name}
% Text
